---
title: "Homework 5"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

# 1

## a
```{r}
data <- read.csv('Vocab.csv')

X <- cbind(data$education, rep(1, length(data$education))) 
Y <- data$vocabulary #scores
#X t Xw = X t y
```

## b
```{r}
xtx <- t(X) %*% X
xty <- t(X) %*% Y

W <- solve(xtx) %*% xty
a <- W[1]
b <- W[2]
a
b
```

## c
```{r}
k <- X%*%W

plot(X[,1], X%*%W)
```
as observed from the plot, people with more education do tend to have larger vocabularies.

# d

```{r}
change_in_marks  <- 1 * W[1]
change_in_marks
```
With unit increase in year of education, there will be 0.33 increase in voculabulary score.


# 2
## a
```{r}
aus_data <- read.csv("ais.csv",stringsAsFactors=FALSE, sep=",")

Y <- aus_data$rcc
X <- aus_data[,3:12]

xtx <- t(X) %*% as.matrix(X)
xty <- t(X) %*% Y

W <- solve(xtx) %*% xty
W
```

## b

```{r}
y_pred <- as.matrix(X) %*% W

error <- (Y - y_pred)
error_square <- error^2
sum_error <- sum(error_square)
sum_error
```

## c

```{r}
lis_errors <- c()
for(i in 1:ncol(X)){
    new_data <- X
    new_data <- new_data[-i] 

    W <- solve(t(new_data) %*% as.matrix(new_data)) %*% (t(new_data) %*% Y)
    y_pred <- as.matrix(new_data) %*% W

    error <- (Y - y_pred)
    error_square <- error^2
    sum_error <- sum(error_square)
    print(sum_error)
    lis_errors <- c(lis_errors, sum_error)
}
print(lis_errors)
```
Variable ‘hc’ ommision causes the greatest increase in sse. Thus 'hc' is the most important variable.

# 3

## a
```{r}
data(nottem)
y <- nottem
n <- length(y)
x <- 1:n

plot(x,y,type="b")
```

## b
```{r}
x_cos <- cos((2*pi*x)/12)
x_sin <- sin((2*pi*x)/12)
for_c <- rep(1,n)

sine.cosine.x <- cbind(x_cos, x_sin, for_c)

w <- solve(t(sine.cosine.x) %*% sine.cosine.x) %*% (t(sine.cosine.x) %*% y)
w
y_pred <- sine.cosine.x %*% w

plot(x,y,type="b")
lines(x,y_pred,type="b",col="red")
```

## c

```{r}
sine.cosine.x_new <- cbind(sine.cosine.x, x)
w_new <- solve(t(sine.cosine.x_new) %*% as.matrix(sine.cosine.x_new)) %*% (t(sine.cosine.x_new) %*% y)
y_pred_d <- sine.cosine.x_new %*% w_new

plot(x,y,type="b")
lines(y_pred_d, type = 'b',col='red')

```
As observd form the plot, it can be seen that the sales have increased slightly. 
Also, the coeficients are positive for x.

# 4

## a

```{r}
X1 <- read.table("pred1.dat")
Y1 <- read.table('resp1.dat')

X1_1 <- X1[1:nrow(X1)/2, ]
Y1_1 <- Y1[1:nrow(X1)/2, ]
W1 <- solve(t(X1_1) %*% as.matrix(X1_1)) %*% (t(X1_1) %*% Y1_1)
W1

X2 <- read.table("pred2.dat")
Y2 <- read.table('resp2.dat')

X2_1 <- X2[1:nrow(X2)/2, ]
Y2_1 <- Y2[1:nrow(X2)/2, ]
W2 <- solve(t(X2_1) %*% as.matrix(X2_1)) %*% (t(X2_1) %*% Y2_1)
W2
```

## b

```{r}
X1_2 <- X1[(nrow(X1)/2+1):nrow(X1), ]
Y1_2 <- Y1[(nrow(X1)/2+1):nrow(X1), ]
SSE_1 <- sum(((as.matrix(X1_2) %*% as.matrix(W1)) - Y1_2)^2)
SSE_1

X2_2 <- X2[((nrow(X2)/2) + 1):nrow(X2), ]
Y2_2 <- Y2[((nrow(X2)/2) + 1):nrow(X2), ]
SSE_2 <- sum(((as.matrix(X2_2) %*% as.matrix(W2)) - Y2_2)^2)
SSE_2
```

# 5

## a

```{r}
X1 <- read.table("pred1.dat")
Y1 <- read.table('resp1.dat')

X1_1 <- X1[1:(nrow(X1)/2), ]
Y1_1 <- Y1[1:(nrow(X1)/2), ]

X1_2 <- X1[((nrow(X1)/2)+1):nrow(X1), ]
Y1_2 <- Y1[((nrow(X1)/2)+1):nrow(Y1), ]

SSE <- rep(0, ncol(X1_1))
for(elem in 1:length(X1_1)){
    column <- X1_1[,elem]
    W <- solve(t(column) %*% as.matrix(column)) %*% (t(column) %*% Y1_1)
    new <- column %*% W
    SSE_1 <- sum(((as.matrix(column) %*% as.matrix(W)) - Y1_1)^2)
    SSE[elem] <- SSE_1    
}
index_single <- which.min(SSE)
SSE[index_single]
index_single

SSE_second <- rep(0, ncol(X1_1))
for(i in 1:ncol(X1_1)){
    if(i == index_single){
        SSE_second[i] = Inf
    }

    else{
        X <- X1_1[,c(i,index_single)]
        w <- solve(t(X) %*% as.matrix(X)) %*% (t(X) %*% Y1_1)
        y_hat <- as.matrix(X) %*% as.matrix(w)
        SSE_ <- sum((y_hat - Y1_1)^2)
        SSE_second[i] <- SSE_    
    }
}
index_single_sec <- which.min(SSE_second)
SSE_second[index_single_sec]
index_single_sec

SSE_third <- rep(0, ncol(X1_1))
for(i in 1:ncol(X1_1)){
    if(i == index_single | i == index_single_sec){
        SSE_third[i] = Inf
    }

    else{
        X <- X1_1[,c(i,index_single, index_single_sec)]
        w <- solve(t(X) %*% as.matrix(X)) %*% (t(X) %*% Y1_1)
        y_hat <- as.matrix(X) %*% as.matrix(w)
        SSE_ <- sum((y_hat - Y1_1)^2)
        SSE_third[i] <- SSE_    
    }
}
index_single_third <- which.min(SSE_third)
SSE_third[index_single_third]
index_single_third
```

## b

```{r}
X_new <- X1_1[,c(index_single, index_single_sec, index_single_third)]
Y_new <- Y1_1

W_new <- solve(t(X_new) %*% as.matrix(X_new)) %*% (t(as.matrix(X_new)) %*% as.matrix(Y_new))
y_hat_new <- as.matrix(X_new) %*% as.matrix(W_new)

SSE_new <- sum((y_hat_new - Y_new)^2)
SSE_new

W_ll <- solve(t(X1_1) %*% as.matrix(X1_1)) %*% (t(X1_1) %*% Y1_1)
y_all <- as.matrix(X1_1) %*% as.matrix(W_ll)

SSE_all <- sum((y_all - Y1_1)^2)
SSE_all
```

# 6

## a

```{r}
X2 <- read.table("pred2.dat")
Y2 <- read.table('resp2.dat')

X2_1 <- X2[1:nrow(X2)/2, ]
Y2_1 <- Y2[1:nrow(X2)/2, ]

X2_2 <- X2[((nrow(X2)/2) + 1):nrow(X2), ]
Y2_2 <- Y2[((nrow(X2)/2) + 1):nrow(X2), ]
lambda <- 20*diag(length(X2_1))
W2_r <- solve(t(X2_1) %*% as.matrix(X2_1) + lambda) %*% (t(X2_1) %*% Y2_1)
y_hat_r <- as.matrix(X2_2) %*% as.matrix(W2_r)
```

## b

```{r}
SSE_r <- sum((y_hat_r - Y2_2)^2)
SSE_r

W2_n <- solve(t(X2_1) %*% as.matrix(X2_1)) %*% (t(X2_1) %*% Y2_1)
y_hat_n <- as.matrix(X2_2) %*% as.matrix(W2_n)
SSE_n <- sum((y_hat_n - Y2_2)^2)
SSE_n
SSE_r
```

## c

```{r}
SSE_list <- c()
for(lambd in 1:40){
    
    # if(! lambd %% 1){
        W <- solve(t(X2_1) %*% as.matrix(X2_1) + lambd*diag(length(X2_1))) %*% (t(X2_1) %*% Y2_1)
        y_hat <- as.matrix(X2_2) %*% as.matrix(W)
        SSE <- sum((y_hat - Y2_2)^2)
        # print(lambd)
        # print(SSE)
        SSE_list <- c( SSE_list,SSE)
    # }

}
length(SSE_list)
plot(SSE_list)
```

# 7

```{r}
data <- read.table("time_series.dat")
n <- nrow(data)
fet1 <- data[1:(n-2),]
fet2 <- data[2:(n-1),]
Y <- data[3:n,]

X <- cbind(fet1, fet2)

w <- solve(t(X) %*% as.matrix(X)) %*% (t(X) %*% Y)
y_pred <- X %*% w

diff <- y_pred - Y
mean(Y - X %*% w)
var(Y - X %*% w)
```