---
title: "Take home exam"
---

# 1

- H0 : number of misprints follow poisson
- H1 : number of misprints do not follow poisson

```{r}
num_mis_print <- c(0:10)
news_pages <- c(9, 23, 40, 30, 31, 26, 19, 10, 5, 4, 3)

mis_print_total <- sum(num_mis_print*news_pages)
news_pages_total <- sum(news_pages)
ave <- mis_print_total/news_pages_total

cats <- news_pages_total*dpois(0:20, ave)
data.frame(0:20, expected=round(cats, 1))

expected <- rep(0, 9)
expected[1:8] <- news_pages_total*dpois(0:7, ave)
expected[9] <- news_pages_total*(1 - ppois(7, ave))
observed <- c(9, 23, 40, 30, 31, 26, 19, 10, 12)

statsL <- 2*sum(observed*log(observed/expected))
statsL
1 - pchisq(statsL, df=7)

statsX  <- sum((observed-expected)^2/expected)
statsX
1 - pchisq(statsX, df=7)
```

If we consider the log likelihood, the calculated probability is more than
the alpha value. Thus, we fail to reject the null hypothesis. The data is consistent 
with the null hypothesis and the number of misprtints do follow a poisson distribution.


If we consider the other test, the calculated probability is less than
the alpha value. Thus, we  reject the null hypothesis. The data is not consistent 
with the null hypothesis and the number of misprtints don't follow a poisson distribution.

# 2

## a

```{r}
exam <- read.table('examanxiety.txt', header=TRUE)

X <- exam$Anxiety
Y <- exam$Exam

corelation <- cor(X,Y)
sy <- sd(Y)
sx <- sd(X)

b <- corelation*(sy/sx)
a <- mean(Y) - b*mean(X)

pred_Y <- a + b*(X)
pred_Y

plot(X,Y)
abline(a,b, col='red')
```

## b

We assumed that the given varialbles have a linear relationship.
We want to predict the exam score according to the anxiety. 

- The Slope is basically the value of increase in exam score for unit increase in anxiety.
- Here the the exam score decreases by around -0.73 for unit increase in anxiety.
- On the other hand the 'a' value is the exam score if the particular candidate has 
zero anxiety. 
- Here the intercept value is around 111.244. The exam score is out of 100 not 111, but I don't think there's anyone who will have zero anxiety!!